% /*@@
%   @file      Infrastructure.tex
%   @date      27 Jan 1999
%   @author    Tom Goodale, Gabrielle Allen, Gerd Lanferman, Thomas Radke
%   @desc
%              Infrastructure thorn writer's guide for the Cactus User's Guide
%   @enddesc
%   @version   $Header$
% @@*/

\renewcommand{\thepage}{\Alph{part}\arabic{page}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\chapter{Infrastructure Thorns}
\label{chap:infrastructure}

\begin{itemize}
 \item{} Concepts and terminology (Overloading and registration of functions)
 \item{} The cGH structure --- what it is and how to use it
 \item{} Extending the cGH structure
 \item{} Querying group and variable information
 \item{} Providing an I/O layer
 \item{} Providing a communication layer
 \item{} Providing a reduction operator
 \item{} Providing an interpolation operator
 \item{} Overloadable functions
\end{itemize}

\section{Concepts and Terminology}
\label{chap:cote}

\subsection{Overloading and Registration}

The flesh defines a core API which guarantees the presence of a set of
functions.  Although the flesh guarantees the presence of these functions,
they can be provided by thorns.  Thorns do this by either the \textit{overloading}
or the \textit{registration} of functions.

\subsubsection{Overloading}

Some functions can only be provided by one thorn.  The first thorn to
\textit{overload} this function succeeds, and any later attempt to overload
the function fails.  For each overloadable function, there is a function
with a name something like {\tt CCTK\_Overload...} which is passed the
function pointer.

\subsubsection{Registration}

Some functions may be provided by several thorns.  The thorns \textit{register}
their function with the flesh, and when the flesh-provided function is called,
the flesh calls all the registered functions.

\subsection{GH Extensions}

A GH extension is a way to associate data with each cGH.  This data should be data
that is required to be associated with a particular GH by a thorn.

Each GH extension is given a unique handle.

\subsection{I/O Methods}

An I/O method is a distinct way to output data.  Each I/O method has a unique name,
and the flesh-provided I/O functions operate on all registered I/O methods.

\section{GH Extensions}

A GH extension is created by calling {\tt CCTK\_RegisterGHExtension}, with the
name of the extension.  This returns a unique handle that identifies the extension.
(This handle can be retrieved at any time by a call to {\tt CCTK\_GHExtensionHandle}.)

Associated with a GH extension are three functions

\begin{Lentry}
\item[{\tt SetupGH}]
this is used to actually create the data structure holding the extension.  It
is called when a new cGH is created.
\item[{\tt InitGH}]
this is used to initialise the extension.  It is called after the scheduler has
been initialised on the cGH.
\item[{\tt ScheduleTraverseGH}]
this is called whenever the schedule tree is due to be traversed on the GH.  It
should initialise the data on the cGH and the call {\tt CCTK\_ScheduleTraverse} to traverse
the schedule tree.
\end{Lentry}

\section{Overloadable and Registerable Functions in Main}

 \begin{tabular}{|l|l|}
   \hline  {\bf Function} & {\bf Default} \\
   \hline {\t CCTK\_Initialise}           &\\
   \hline {\t CCTK\_Evolve}               &\\
   \hline {\t CCTK\_Shutdown}             &\\
   \hline
 \end{tabular}

\section{Overloadable and Registerable Functions in Comm}

  \begin{tabular}{|l|l|}
   \hline {\bf Function} & {\bf Default}  \\
   \hline {\t CCTK\_SyncGroup}           &\\
   \hline {\t CCTK\_SyncGroupsByDirI}    &\\
   \hline {\t CCTK\_EnableGroupStorage}  &\\
   \hline {\t CCTK\_DisableGroupStorage} &\\
   \hline {\t CCTK\_QueryMaxTimeLevels}  &\\
   \hline {\t CCTK\_EnableGroupComm}     &\\
   \hline {\t CCTK\_DisableGroupComm}    &\\
   \hline {\t CCTK\_Barrier}             &\\
   \hline {\t CCTK\_Reduce}              &\\
   \hline {\t CCTK\_Interp}              &\\
   \hline {\t CCTK\_ParallelInit}        &\\
   \hline
  \end{tabular}

\section{Overloadable and Registerable Functions in I/O}

 \begin{tabular}{|l|l|}
   \hline {\bf Function} & {\bf Default}   \\
   \hline {\t CCTK\_OutputGH}            & \\
   \hline {\t CCTK\_OutputVarAsByMethod} & \\
   \hline
 \end{tabular}

\section{Drivers}

The flesh does not know about memory allocation for grid variables, about how
to communicate data when synchronisation is called for, or about
multiple patches or adaptive mesh refinement.  All this is the job of a driver.

This chapter describes how to add a driver to your code.

\subsection{Anatomy}

A driver consists of a Startup routine which creates a GH extension, 
registers its associated functions, and overloads the communication functions.
It may optionally register interpolation, reduction, and I/O methods.

A driver may also overload the default Initialisation and Evolution routines,
although a simple unigrid evolver is supplied in the flesh.

\subsection{Startup}

A driver consists of a GH extension, and the following overloaded
functions.

\begin{enumerate}
\item{} {\tt CCTK\_EnableGroupStorage}
\item{} {\tt CCTK\_DisableGroupStorage}
\item{} {\tt CCTK\_QueryMaxTimeLevels}
\item{} {\tt CCTK\_ArrayGroupSizeB}
\item{} {\tt CCTK\_QueryGroupStorageB}
\item{} {\tt CCTK\_SyncGroup}
\item{} {\tt CCTK\_SyncGroupsByDirI}
\item{} {\tt CCTK\_EnableGroupComm}
\item{} {\tt CCTK\_DisableGroupComm}
\item{} {\tt CCTK\_Barrier}
\item{} {\tt CCTK\_OverloadParallelInit}
\item{} {\tt CCTK\_OverloadExit}
\item{} {\tt CCTK\_OverloadAbort}
\item{} {\tt CCTK\_OverloadMyProc}
\item{} {\tt CCTK\_OverloadnProcs}
\end{enumerate}

The overloadable function {\tt CCTK\_SyncGroup} is deprecated, a driver should
instead provide a routine to overload the more general function
{\tt CCTK\_SyncGroupsByDirI}.

\subsection{The GH Extension}

The GH extension is where the driver stores all its grid-dependent information.
This is stuff like any data associated with a grid variable (e.g.\ storage and
communication state), how many grids if it is AMR, ...  It is very difficult to
describe in general, but one simple example might be

\begin{verbatim}

struct SimpleExtension
{
  /* The data associated with each variable */
  /* data[var][timelevel][ijk]                */
  void ***data
} ;

\end{verbatim}

with a {\tt SetupGH} routine like

\begin{verbatim}

struct SimpleExtension *SimpleSetupGH(tFleshConfig *config, int conv_level, cGH *GH)
{
   struct SimpleExtension *extension;

   extension = NULL;

   if(conv_level < max_conv_level)
   {
      /* Create the extension */
      extension = malloc(sizeof(struct SimpleExtension));

      /* Allocate data for all the variables */
      extension->data = malloc(num_vars*sizeof(void**));

      for(var = 0 ; var < num_vars; var++)
      {
        /* Allocate the memory for the time levels */
        extension->data[var] = malloc(num_var_time_levels*sizeof(void *));

        for(time_level = 0; time_level < num_var_time_level; time_level++)
        {
          /* Initialise the data to NULL */
          extension->data[var][time_level] = NULL;
        }
      }
    }

   return extension;
}

\end{verbatim}

Basically, what this example is doing is preparing a data array for use.  The
function can query the flesh for information on every variable.  Note that
scalars should always have memory actually assigned to them.

An {\tt InitGH} function isn't strictly necessary, and in this case, it could
just be a dummy function.

The {\tt ScheduleTraverseGH} function needs to fill out the cGH data, and
then call {\tt CCTK\_ScheduleTraverse} to have the functions scheduled at
that point executed on the grid

\begin{verbatim}

int SimpleScheduleTraverseGH(cGH *GH, const char *where)
{
  int retcode;
  int  var;
  int  gtype;
  int  ntimelevels;
  int  level;
  int  idir;

  extension = (struct SimpleExtension *)GH->extensions[SimpleExtension];

  for (idir=0;idir<GH->cctk_dim;idir++)
  {
    GH->cctk_levfac[idir] = 1;
    GH->cctk_nghostzones[idir] = extension->nghostzones[idir];
    GH->cctk_lsh[idir]         = extension->lnsize[idir];
    GH->cctk_gsh[idir]         = extension->nsize[idir];
    GH->cctk_bbox[2*idir]      = extension->lb[extension->myproc][idir] == 0;
    GH->cctk_bbox[2*idir+1]    = extension->ub[extension->myproc][idir]
                              == extension->nsize[idir]-1;
    GH->cctk_lbnd[idir]        = extension->lb[extension->myproc][idir];
    GH->cctk_ubnd[idir]        = extension->ub[extension->myproc][idir];

  }

  for(var = 0; var < extension->nvariables; var++)
  {
    gtype = CCTK_GroupTypeFromVarI(var);
    ntimelevels = CCTK_MaxTimeLevelsVI(var);

    for(level = 0; level < ntimelevels; level++)
    {
      switch(gtype)
      {
        case CCTK_SCALAR :
          GH->data[var][level] = extension->variables[var][level];
          break;
        case CCTK_GF     :
          GH->data[var][level] =
            ((pGF ***)(extension->variables))[var][level]->data;
          break;
        case CCTK_ARRAY :
          GH->data[var][level] =
            ((pGA ***)(extension->variables))[var][level]->data;
          break;
        default:
          CCTK_WARN(CCTK_WARN_ALERT,"Unknown group type in SimpleScheduleTraverse");
      }
    }
  }

  retcode = CCTK_ScheduleTraverse(where, GH, NULL);

  return retcode;

}

\end{verbatim}

The third argument to {\tt CCTK\_ScheduleTraverse} is actually a function
which will be called by the scheduler when it wants to call a function
scheduled by a thorn.  This function is given some information about
the function to call, and is an alternative place where the cGH can be setup.

This function is optional, but a simple implementation might be

\begin{verbatim}

int SimpleCallFunction(void *function,
                       cFunctionData *fdata,
                       void *data)
{
  void (*standardfunc)(void *);

  int (*noargsfunc)(void);

  switch(fdata->type)
  {
    case FunctionNoArgs:
      noargsfunc = (int (*)(void))function;
      noargsfunc();
      break;
    case FunctionStandard:
      switch(fdata->language)
      {
        case LangC:
          standardfunc = (void (*)(void *))function;
          standardfunc(data);
          break;
        case LangFortran:
          fdata->FortranCaller(data, function);
          break;
        default :
          CCTK_WARN(CCTK_WARN_ALERT, "Unknown language.");
      }
      break;
    default :
      CCTK_WARN(CCTK_WARN_ALERT, "Unknown function type.");
  }

  /* Return 0, meaning didn't synchronise */
  return 0;
}

\end{verbatim}

The return code of the function signifies whether or not the function
synchronised the groups in this functions synchronisation list of not.

The flesh will synchronise them if the function returns false.

Providing this function is probably the easiest way to do multi-patch or
AMR drivers.

\subsection{Memory Functions}

These consist of
\begin{enumerate}
\item{} {\tt CCTK\_EnableGroupStorage}
\item{} {\tt CCTK\_DisableGroupStorage}
\item{} {\tt CCTK\_QueryMaxTimeLevels}
\item{} {\tt CCTK\_QueryGroupStorageB}
\item{} {\tt CCTK\_ArrayGroupSizeB}
\end{enumerate}

\subsubsection{En/Disable Group Storage}

These are responsible for switching the memory for all variables
in a group on or off.  They should return the former state, e.g.\
if the group already has storage assigned, they should return 1.

In our simple example above, the enabling routine would look
something like

\begin{verbatim}

int SimpleEnableGroupStorage(cGH *GH, const char *groupname)
{

  extension = (struct SimpleExtension *)GH->extensions[SimpleExtension];

  if(extension->data[first][0][0] == NULL)
  {
    for(var = first; var <= last; var++)
    {
      allocate memory for all time levels;
    }
    retcode = 0;
  }
  else
  {
    retcode = 1;
  }

  return retcode;
}

\end{verbatim}

The disable function is basically the reverse of the enable one.

The {\tt CCTK\_QueryMaxTimeLevels} function returns the maximum number of
timelevels ever activated for a given group ie. the size of the {\tt data}
member of {\tt cGH}.

The {\tt QueryGroupStorage} function basically returns true or false if
there is storage for the group, and the {\tt ArrayGroupSize} returns the
size of the grid function or array group in a particular direction.

\subsubsection{En/Disable Group Comm}

These are the communication analogues to the storage functions.  Basically,
they flag that communication is to be done on that group or not, and may
initialise data structures for the communication.


\section{I/O Methods}
\label{chap:io_methods}
%
The flesh by itself does not provide output for grid variables or other data.
Instead, it provides a mechanism for thorns to register their own
routines as I/O methods, and to invoke these I/O methods by either the
flesh scheduler or by other thorn routines.

This chapter explains how to implement your own I/O methods and register
them with the flesh.
%
\subsection{I/O Method Registration}
%
All I/O methods have to be registered with the flesh before they can be used.

The flesh I/O registration API provides a routine {\t CCTK\_RegisterIOMethod()}
to create a handle for a new I/O method which is identified by its name
(this name must be unique for all I/O methods).
With such a handle, a thorn can then register a set of functions (using the
{\t CCTK\_RegisterIOMethod*()} routines from the flesh) for doing
periodic, triggered, and/or unconditional output.

The following example shows how a thorn would register an I/O method,
{\tt SimpleIO}, with routines to provide all these different types of output.
%
\begin{verbatim}
  void SimpleIO_Startup (void)
  {
    int handle = CCTK_RegisterIOMethod ("SimpleIO");
    if (handle >= 0)
    {
      CCTK_RegisterIOMethodOutputGH (handle, SimpleIO_OutputGH);

      CCTK_RegisterIOMethodTimeToOutput (handle, SimpleIO_TimeToOutput);
      CCTK_RegisterIOMethodTriggerOutput (handle, SimpleIO_TriggerOutput);

      CCTK_RegisterIOMethodOutputVarAs (handle, SimpleIO_OutputVarAs);
    }
  }
\end{verbatim}
%
%
\subsection{Periodic Output of Grid Variables}
%
The flesh scheduler will automatically call {\t CCTK\_OutputGH()} at every
iteration after the {\tt CCTK\_ANALYSIS} time bin. This function loops over all
I/O methods and calls their routines registered as {\t OutputGH()} (see also Section
\ref{subsec:schedule_ccl}).
%
\begin{alltt}
  int SimpleIO_OutputGH (const cGH *\var{GH});
\end{alltt}
%
The {\t OutputGH()} routine itself should loop over all variables living on the
\texttt{GH} grid hierarchy, and do all necessary output if requested
(this is usually determined by I/O parameter settings).

As its return code, it should pass back the number of variables which were output
at the current iteration, or zero if no output was done by this I/O method.
%
%
\subsection{Triggered Output of Grid Variables}
%
Besides the periodic output at every so many iterations using {\t OutputGH()},
analysis and output of grid variables can also be done via triggers. For this,
a {\t TimeToOutput()} routine is registered with an I/O method.
This routine will be called by the flesh scheduler at every iteration before
{\tt CCTK\_ANALYSIS} with the triggering variable(s) as defined in the schedule
block for all {\tt CCTK\_ANALYSIS} routines (see Section \ref{scheduling:schedule_block}).

If the {\t TimeToOutput()} routine decides that it is now time to do output, the
flesh scheduler will invoke the corresponding analysis routine and also request
output of the triggering variable(s) using {\t TriggerOutput()}.
%
\begin{alltt}
  int SimpleIO_TimeToOutput (const cGH *\var{GH}, int \var{varindex});
  int SimpleIO_TriggerOutput (const cGH *\var{GH}, int \var{varindex});
\end{alltt}
%
Both routines get passed the index of a possible triggering grid variable.

{\t TimeToOutput()} should return a non-zero value if analysis and output
for \texttt{\var{varindex}} should take place at the current iteration, and zero
otherwise.

{\t TriggerOutput()} should return zero for successful output of variable
\texttt{\var{varindex}}, and a negative value in case of an error.
%
%
\subsection{Unconditional Output of Grid Variables}

An I/O method's {\t OutputVarAs()} routine is supposed to do output for a
specific grid variable if ever possible. It will be invoked by the flesh I/O API
routines {\t CCTK\_OutputVar*()} for unconditional, non-triggered output of
grid variables (see also Section \ref{sec:io}).

A function registered as an \texttt{OutputVarAs()} routine has the following
prototype:
%
\begin{alltt}
  int SimpleIO_OutputVarAs (const cGH *\var{GH}, const char *\var{varname}, const char *\var{alias});
\end{alltt}
%
The variable to output, \texttt{\var{varname}}, is given by its full name.
The full name may have appended an optional I/O options string enclosed in
curly braces (with no space between the full name and the opening curly brace).
In addition to that, an \texttt{\var{alias}} string can be passed which then
serves the purpose of constructing a unique name for the output file.

The {\t OutputVarAs()} routine should return zero if output for
\texttt{\var{varname}}
was done successfully, or a negative error code otherwise.


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Checkpointing/Recovery Methods}
\label{chap:cp_recovery_methods}

Like for I/O methods, checkpointing/recovery functionality must be implemented
by thorns. The flesh only provides specific time bins at which the scheduler
will call thorns' routines, in order to perform checkpointing and/or recovery.

This chapter explains how to implement checkpointing and recovery methods
in your thorn.  For further information, see the documentation for thorn
\texttt{CactusBase/IOUtil}.

\subsection{Checkpointing Invocation}

Thorns register their checkpointing routines at {\t CCTK\_CPINITIAL} and/or
{\t CCTK\_CHECKPOINT} and/or {\t CCTK\_TERMINATE}. These time bins are
scheduled right after all initial data has been set up, after every
evolution timestep, and after the last time step of a simulation,
respectively. (See Section \ref{subsec:schedule_ccl} for a description of all timebins).
Depending on
parameter settings, the checkpoint routines decide whether to write an initial
data checkpoint, and when to write an evolution checkpoint.

Each checkpoint routine should save all information to persistent storage, which
is necessary to restart the simulation at a later time from exactly the same
state. Such information would include
%
\begin{itemize}
  \item the current settings of all parameters
  \item the contents of all grid variables which have global storage assigned
    and are not tagged with {\tt checkpoint="no"} (see also Section \ref{subsec:Appendix.interface-variables} on page \pageref{subsec:Appendix.interface-variables} for a list of possible tags)\\
    Note that grid variables should be synced before writing them to disk.
  \item other relevant information such as the current iteration number and
    physical time, the number of processors, etc.
\end{itemize}


\subsection{Recovery Invocation}

Recovering from a checkpoint is a two-phase operation for which the flesh
provides distinct timebins for recovery routines to be scheduled at:
%
\begin{Lentry}
  \item[{\t CCTK\_RECOVER\_PARAMETERS}] This time bin is executed before
    {\t CCTK\_STARTUP}, in which the parameter file is parsed. From these parameter
    settings, the recovery routines should decide whether recovery was
    requested, and if so, restore all parameters from the checkpoint file and
    overwrite those which aren't steerable.\\
    The flesh loops over all registered recovery routines to find out
    whether recovery was requested. Each recovery routine should, therefore,
    return a positive integer value for successful parameter recovery (causing
    a shortcut of the loop over all registered recovery routines),
    zero for no recovery, or a negative value to flag an error.\\
    If recovery was requested, but no routine could successfully recover
    parameters, the flesh will abort the run with an error message.
    If no routine recovered any parameters, i.e. if all parameter
    recovery routines returned zero, then this indicates that this run
    is not a recovery run.\\
    If parameter recovery was performed successfully, the scheduler will set the
    {\tt recovered} flag which---in combination with the setting of the {\tt
    Cactus::recovery\_mode} parameter---decides whether any thorn routine
    scheduled for {\t CCTK\_INITIAL} and {\t CCTK\_POSTINITIAL} will be called.
    The default is to not execute these initial time bins during recovery, 
    because the initial data will be set up from the checkpoint file during the
    following {\t CCTK\_RECOVER\_VARIABLES} time bin.
  \item[{\t CCTK\_RECOVER\_VARIABLES}]
    Recovery routines scheduled for this time bin are responsible for restoring
    the contents of all grid variables with storage assigned from the
    checkpoint.\\
    Depending on the setting of {\tt Cactus::recovery\_mode}, they should also
    decide how to treat errors in recovering individual grid variables. Strict
    recovery (which is the default) requires all variables to be restored
    successfully (and will stop the code if not), whereas a relaxed mode
    could, e.g. allow for grid variables, which are not found in the checkpoint
    file, to be gracefully ignored during recovery.
\end{Lentry}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Clocks for Timing}
\label{chap:clocks}

To add a Cactus clock, you need to write several functions to provide the
timer functionality (see Section \ref{sec:timers}), and then register these
functions with the flesh as a named clock.

The function pointers are placed in function pointer fields of a 
\texttt{cClockFuncs} structure.  The fields of this structure are are:

\begin{Lentry}

\item[{\t create}] {\t void *(*create)(int)}
\item[{\t destroy}] {\t void (*destroy)(int, void *)}
\item[{\t start}] {\t void (*start)(int, void *)}
\item[{\t stop}] {\t void (*stop)(int, void *)}
\item[{\t reset}] {\t void (*reset)(int, void *)}
\item[{\t get}] {\t void (*get)(int, void *, cTimerVal *)}
\item[{\t set}] {\t void (*set)(int, void *, cTimerVal *)}
\item[{\t n\_vals}] {\t int}

\end{Lentry}

The first \texttt{int} argument of the functions may be used in any way
you see fit.

The \texttt{n\_vals} field holds the number of elements in the 
\texttt{vals} array field of the \texttt{cTimerVal} structure 
used by your timer (usually 1).

The return value of the \texttt{create} function will be a pointer to
a new structure representing your clock.

The second \texttt{void*} argument of all the other functions will be the
pointer returned from the \texttt{create} function.

The \texttt{get} and \texttt{set} functions should write to and read from
(respectively) a structure pointed to by the \texttt{cTimerVal*} argument:

\begin{verbatim}
typedef enum {val_none, val_int, val_long, val_double} cTimerValType;

typedef struct
{
  cTimerValType type;
  const char *heading;
  const char *units;
  union
 {
    int        i;
    long int   l;
    double     d;
  } val;
  double seconds;
  double resolution;
} cTimerVal;
\end{verbatim}

The \texttt{heading} field is the name of the clock, the \texttt{units} 
field  holds a string describing the type held in the \texttt{val} field,
and the \texttt{seconds} field is the time elapsed in seconds.  The 
\texttt{resolution} field is the smallest non-zero difference in values 
of two calls to the timer, in seconds.  For example, it could reflect
that the clock saves the time value internally as an integer value 
representing milliseconds.

To name and register the clock with the flesh, call the function
\begin{alltt}
	CCTK_ClockRegister( "\var{my_clock_name}", &\var{clock_func} ).
\end{alltt}



%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\end{cactuspart}
