
\thispagestyle{empty}

\vspace*{0.1in}

\begin{centering}
{\LARGE \NAMD\ Version \NAMDVER}\\
\bigskip
{\large Authors: \NAMDAUTHORS} \\
\medskip
{\large Theoretical and Computational Biophysics Group, Beckman Institute, University of Illinois.} \\
\bigskip
{\large \copyright 1995-2011 The Board of Trustees of the University of Illinois.
All Rights Reserved} \\
\bigskip
\end{centering}

  \rule{6in}{0.04in}				\\	\vspace{0.25in}

\section*{NAMD Molecular Dynamics Software \\
Non-Exclusive, Non-Commercial Use License}

\subsubsection*{Introduction}

The University of Illinois at Urbana-Champaign has created its molecular
dynamics software, NAMD, developed by the Theoretical and Computational Biophysics 
Group (``TCBG'') at Illinois' Beckman Institute available free of charge for
non-commercial use by individuals, academic or research institutions and
corporations for in-house business purposes only, upon completion and
submission of the online registration form presented when attempting to
download NAMD at the web site {\tt http://www.ks.uiuc.edu/Research/namd/}.

Commercial use of the NAMD software, or derivative works based thereon,
REQUIRES A COMMERCIAL LICENSE. Commercial use includes: 
(1) integration of all or part of the Software into a product for sale, 
lease or license by or on behalf of Licensee to third parties, or 
(2) distribution of the Software to third parties that need it to 
commercialize product sold or licensed by or on behalf of Licensee.  
The University of Illinois will negotiate commercial-use licenses for 
NAMD upon request. These requests can be directed to namd@ks.uiuc.edu

\subsubsection*{Online Download Registration Requirements}

In completing the online registration form presented before downloading
individuals may register in their own name or with their institutional or
corporate affiliations. Registration information must include name, title, 
and e-mail of a person with signature authority to authorize and commit the
individuals, academic or research institution, or corporation as necessary 
to the terms and conditions of the license agreement.

All parts of the information must be understood and agreed to as part of
completing the form. Completion of the form is required before software 
access is granted. Pay particular attention to the authorized requester 
requirements above, and be sure that the form submission is authorized 
by the duly responsible person.

\newpage
\subsubsection*{UNIVERSITY OF ILLINOIS \\
NAMD MOLECULAR DYNAMICS SOFTWARE LICENSE AGREEMENT}

Upon execution of this Agreement by the party identified below (``Licensee''),
The Board of Trustees of the University of Illinois (``Illinois''), on 
behalf of The Theoretical and Computational Biophysics Group (``TCBG'') in the Beckman Institute,
will provide the molecular dynamics software NAMD in Executable 
Code and/or Source Code form (``Software'') to Licensee, subject to 
the following terms and conditions. For purposes of this Agreement, 
Executable Code is the compiled code, which is ready to run on Licensee's 
computer. Source code consists of a set of files which contain the 
actual program commands that are compiled to form the Executable Code.

1. The Software is intellectual property owned by Illinois, and all 
right, title and interest, including copyright, remain with Illinois. 
Illinois grants, and Licensee hereby accepts, a restricted, non-exclusive, 
non-transferable license to use the Software for academic, research 
and internal business purposes only e.g. not for commercial use 
(see Paragraph 7 below), without a fee. Licensee agrees to reproduce 
the copyright notice and other proprietary markings on all copies of 
the Software. Licensee has no right to transfer or sublicense the 
Software to any unauthorized person or entity. However, Licensee does 
have the right to make complimentary works that interoperate with NAMD, 
to freely distribute such complimentary works, and to direct others 
to the TCBG server to obtain copies of NAMD itself.

2. Licensee may, at its own expense, modify the Software to make derivative
works, for its own academic, research, and internal business purposes.
Licensee's distribution of any derivative work is also subject to the same
restrictions on distribution and use limitations that are specified herein
for Illinois' Software. Prior to any such distribution the Licensee shall 
require the recipient of the Licensee's derivative work to first execute a 
license for NAMD with Illinois in accordance with the terms and conditions 
of this Agreement. Any derivative work should be clearly marked and 
renamed to notify users that it is a modified version and not the original 
NAMD code distributed by Illinois.

3. Except as expressly set forth in this Agreement, THIS SOFTWARE IS 
PROVIDED ``AS IS'' AND ILLINOIS MAKES NO REPRESENTATIONS AND EXTENDS NO 
WARRANTIES OF ANY KIND, EITHER EXPRESS OR IMPLIED, INCLUDING BUT NOT 
LIMITED TO WARRANTIES OR MERCHANTABILITY OR FITNESS FOR A PARTICULAR 
PURPOSE, OR THAT THE USE OF THE SOFTWARE WILL NOT INFRINGE ANY PATENT, 
TRADEMARK, OR OTHER RIGHTS. LICENSEE ASSUMES THE ENTIRE RISK AS TO 
THE RESULTS AND PERFORMANCE OF THE SOFTWARE AND/OR ASSOCIATED MATERIALS. 
LICENSEE AGREES THAT UNIVERSITY SHALL NOT BE HELD LIABLE FOR ANY DIRECT, 
INDIRECT, CONSEQUENTIAL, OR INCIDENTAL DAMAGES WITH RESPECT TO ANY 
CLAIM BY LICENSEE OR ANY THIRD PARTY ON ACCOUNT OF OR ARISING FROM 
THIS AGREEMENT OR USE OF THE SOFTWARE AND/OR ASSOCIATED MATERIALS.

4. Licensee understands the Software is proprietary to Illinois. 
Licensee agrees to take all reasonable steps to insure that the Software 
is protected and secured from unauthorized disclosure, use, or release 
and will treat it with at least the same level of care as Licensee would 
use to protect and secure its own proprietary computer programs and/or 
information, but using no less than a reasonable standard of care. 
Licensee agrees to provide the Software only to any other person or 
entity who has registered with Illinois. If licensee is not registering 
as an individual but as an institution or corporation each member of 
the institution or corporation who has access to or uses Software must 
understand and agree to the terms of this license. If Licensee becomes 
aware of any unauthorized licensing, copying or use of the Software, 
Licensee shall promptly notify Illinois in writing. Licensee expressly 
agrees to use the Software only in the manner and for the specific uses 
authorized in this Agreement.

5. By using or copying this Software, Licensee agrees to abide by the
copyright law and all other applicable laws of the U.S. including, but 
not limited to, export control laws and the terms of this license. 
Illinois shall have the right to terminate this license immediately by 
written notice upon Licensee's breach of, or non-compliance with, any 
of its terms. Licensee may be held legally responsible for any copyright 
infringement that is caused or encouraged by its failure to abide by 
the terms of this license. Upon termination, Licensee agrees to destroy 
all copies of the Software in its possession and to verify such 
destruction in writing.

6. The user agrees that any reports or published results obtained 
with the Software will acknowledge its use by the appropriate citation 
as follows: 

\begin{quote}
 NAMD was developed by the Theoretical and Computational Biophysics Group in the 
 Beckman Institute for Advanced Science and Technology at the 
 University of Illinois at Urbana-Champaign.
\end{quote}

Any published work which utilizes NAMD shall include the following reference: 

\begin{quote}
 James C. Phillips, Rosemary Braun, Wei Wang, James Gumbart,
 Emad Tajkhorshid, Elizabeth Villa, Christophe Chipot, Robert D. Skeel,
 Laxmikant Kale, and Klaus Schulten. Scalable molecular dynamics with NAMD.
 {\it Journal of Computational Chemistry}, 26:1781-1802, 2005.
\end{quote}

Electronic documents will include a direct link to the official NAMD page:

\begin{quote}
{\tt http://www.ks.uiuc.edu/Research/namd/}
\end{quote}

One copy of each publication or report will be supplied to Illinois 
at the addresses listed below in Contact Information.

7. Should Licensee wish to make commercial use of the Software, Licensee 
will contact Illinois (namd@ks.uiuc.edu) to negotiate an appropriate 
license for such use. Commercial use includes: (1) integration of all 
or part of the Software into a product for sale, lease or license by or 
on behalf of Licensee to third parties, or (2) distribution of the 
Software to third parties that need it to commercialize product sold or 
licensed by or on behalf of Licensee.

8. Government Rights. Because substantial governmental funds have been 
used in the development of NAMD, any possession, use or sublicense of 
the Software by or to the United States government shall be subject 
to such required restrictions.

9. NAMD is being distributed as a research and teaching tool and as such, 
TCBG encourages contributions from users of the code that might, at 
Illinois' sole discretion, be used or incorporated to make the basic 
operating framework of the Software a more stable, flexible, and/or 
useful product. Licensees that wish to contribute their code to become 
an internal portion of the Software may be required to sign an 
``Agreement Regarding Contributory Code for NAMD Software''
before Illinois can accept it (contact namd@ks.uiuc.edu for a copy).

\newpage
\subsubsection*{Contact Information}

The best contact path for licensing issues is by e-mail to namd@ks.uiuc.edu
or send correspondence to:
\begin{verse}
                             NAMD Team\\
                             Theoretical and Computational Biophysics Group\\
                             Beckman Institute\\
                             University of Illinois\\
                             405 North Mathews MC-251\\
                             Urbana, Illinois 61801 USA
\end{verse}


